### 2014-10-22 TransitionsUseful.tex
few corrections
stuff

%\begin{figure}
%\small
%\centering
%\begin{boxedminipage}[c]{.48\textwidth}
%%\centering
%\inference
%[Halt]
%{\theta(i) = (s, \pc) & \dec(s,\pc) = H}
%{\io{SC}{\theta,m}{H}{}}
%
%\inference
%[NonMem]
%{\theta(i) = (s, \pc) & \dec(s,\pc) = (\nm, x) \\ \exec(s, \pc, (\nm, x)) = (\pc', s')}
%{\io{SC}{\theta,m}{\theta[i \coloneqq (s', \pc')], m}{}}
%
%\inference
%[Load]
%{\theta(i) = (s,\pc) & \dec(s,\pc) = (\ld, x, a) \\ \exec(s, \pc, (\ld, x, m(a))) = (\pc', s')}
%{\io{SC}{\theta,m}{\theta[i\coloneqq(s',\pc')],m}{}}
%
%\inference
%[Store]
%{\theta(i) = (s,\pc) & \dec(s,\pc) = (\st, a, v) \\ \exec(s, \pc, (\st)) = (\pc', s')}
%{\io{SC}{\theta,m}{\theta[i\coloneqq(s',\pc')],m[a\coloneqq v]}{}}
%\end{boxedminipage}
%\caption{LTS for sequential consistency with $n$ simple processors}
%\label{Ref}
%\end{figure}
%
%\section{Decomposing a Multi-Processor System}
%\label{sec:store-atomicity}
%
%\begin{figure}
%\centering
%\includegraphics[scale=.43]{zoom}
%\caption{Components of a multi-processors system}
%\label{zoom}
%\end{figure}

@@

\begin{figure}
\small
\centering
\begin{boxedminipage}[c]{.48\textwidth}
\inference
[Load]
{}
{\io{$M_m$}{m}{m}{(\ld\req,a),(\ld\resp,m(a)),\_}}

\inference
[Store]
{}
{\io{$M_m$}{m}{m[a\coloneqq v]}{(\st, a, v)}}

@@

So, we focus on this opportunity to simplify proof decomposition.  We
will prove that our main memory component satisfies an intuitive
\emph{store atomicity} property.  Despite the potential red herring of
the word ``atomicity,'' this specification is appropriate even for
implementation of weaker memory models than SC.  Store atomicity can
be understood via the operational semantics of Figure \ref{$M_m$},
describing an LTS that receives load and store requests ($\ld$ and
$\st$) from processors and sends back load responses ($\ld\resp$).

@@

The interface and formal contract of a reorder buffer are more
involved.  We write $\rob$ as a metavariable for reorder buffer
state.  The associated operations are:
\begin{itemize}
\item $\phi$, the state of an empty buffer.
\item $\add(\pc, \rob)$, which appends the program instruction at location $\pc$ to the list of instructions that the buffer is allowed to consider executing.
\item $\get\spc\ld(\rob)$, which returns an optional speculative load to issue
\item $\compute(\rob)$, which models a
step of computation inside the buffer. For instance, it invokes the $\dec$
and $\exec$ functions (as defined for SC) internally to obtain the next program
counter, state, \etc{} (only the speculative states are updated, not the actual states).
\item $\upd\ld(\rob,t,v)$, which informs the buffer that the memory
has returned result value $v$ for the speculative load with tag $t$.
\item $\commit(\rob)$, which returns the next instruction in serial
program order, if we have accumulated enough memory responses to execute it
accurately, or returns
$\epsilon$ otherwise.  When $\commit$ returns an instruction, it also
returns the associated program counter plus the next program counter
we would advance to afterward.
 Furthermore, the instruction is
extended with any relevant response from memory (used only for load
instructions, obtained through $\upd\ld$) and with the new architectural state (\eg{} register
file) after execution.
\item $\retire(\rob)$, which informs the buffer that its $\commit$
instruction was executed successfully, so it is time to move on to the
next instruction.
\end{itemize}

@@

\inference[Compute]
{\get\spc\ld(\rob) = \epsilon}
{\ioe{\text{P$_\text{so}$}}{s,\pc,\rob,\ppc}{s,\pc,\compute(\rob),\ppc}}

\inference[SpecLd]
{\get\spc\ld(\rob) = (\spc\ld,t,a)}
{\io{\text{P$_\text{so}$}}{s,\pc,\rob,\ppc}{s,\pc,\upd\ld(\compute(\rob),t,v),\ppc}
{(\ld\req, a), (\ld\resp, v), \Spec}}

@@

\inference[Load]
{\s(c,a) \geq S}
{\io{$M_c$}
{\dt, \ch, \s, \dst, \wt, \dwt}
{\dt, \ch, \s, \dst, \wt, \dwt}{c,((\ld\req, a),(\ld\resp,v)), \_}}
\vspace{.1in}
\inference[Store]
{s(c,a) \geq M}
{\io{$M_c$}
{\dt, \ch, \s, \dst, \wt, \dwt}
{\dt[(c, a) \coloneqq v], \ch, \s, \dst, \wt, \dwt}{c,(\st,a,v)}}

>>>>

%\begin{figure}
%\small
%\centering
%\begin{boxedminipage}[c]{.48\textwidth}
%%\centering
%\inference
%[Halt]
%{\theta(i) = (s, \pc) & \dec(s,\pc) = H}
%{\io{SC}{\theta,m}{H}{}}
%
%\inference
%[NonMem]
%{\theta(i) = (s, \pc) & \dec(s,\pc) = (\nm, x) \\ \exec(s, \pc, (\nm, x)) = (\pc', s')}
%{\io{SC}{\theta,m}{\theta[i \coloneqq (s', \pc')], m}{}}
%
%\inference
%[Load]
%{\theta(i) = (s,\pc) & \dec(s,\pc) = (\ld, x, a) \\ \exec(s, \pc, (\ld, x, m(a))) = (\pc', s')}
%{\io{SC}{\theta,m}{\theta[i\coloneqq(s',\pc')],m}{}}
%
%\inference
%[Store]
%{\theta(i) = (s,\pc) & \dec(s,\pc) = (\st, a, v) \\ \exec(s, \pc, (\st)) = (\pc', s')}
%{\io{SC}{\theta,m}{\theta[i\coloneqq(s',\pc')],m[a\coloneqq v]}{}}
%\end{boxedminipage}
%\caption{LTS for sequential consistency with $n$ simple processors}
%\label{Ref}
%\end{figure}
%
\section{Decomposing a Multi-Processor System}
\label{sec:store-atomicity}

\begin{figure}
\centering
\includegraphics[scale=.43]{zoom}
\caption{Components of a multi-processors system}
\label{zoom}
\end{figure}

@@

\begin{figure}
\small
\centering
\begin{boxedminipage}[c]{.48\textwidth}
\inference
[Load]
{}
{\io{$M_m$}{m}{m}{p, (\ld\req,a),(\ld\resp,m(a), \_)}}

\inference
[Store]
{}
{\io{$M_m$}{m}{m[a\coloneqq v]}{p, (\st, a, v)}}

@@

So, we focus on this opportunity to simplify proof decomposition.  We
will prove that our main memory component satisfies an intuitive
\emph{store atomicity} property.  Despite the potential red herring of
the word ``atomicity,'' this specification is appropriate even for
implementation of weaker memory models than SC.  Store atomicity can
be understood via the operational semantics of Figure \ref{$M_m$},
describing an LTS that receives load and store requests ($\ld$ and
$\st$) from processors and sends back load responses ($\ld\resp$);
$p$ is the processor which sends and receives the request.

@@

The interface and formal contract of a reorder buffer are more
involved.  We write $\rob$ as a metavariable for reorder buffer
state.  The associated operations are:
\begin{itemize}
\item $\phi$, the state of an empty buffer.
\item $\add(\pc, \rob)$, which appends the program instruction at location $\pc$ to the list of instructions that the buffer is allowed to consider executing.
\item $\get\spc\ld(\rob)$, which returns a set of speculative loads that can be issued to the memory system
\item $\compute(\rob)$, which models a
step of computation inside the buffer. For instance, it invokes the $\dec$
and $\exec$ functions (as defined for SC) internally to obtain the next program
counter, state, \etc{} (only the speculative states are updated, not the actual states).
\item $\upd\ld(\rob,t,v)$, which informs the buffer that the memory
has returned result value $v$ for the speculative load with tag $t$.
\item $\commit(\rob)$, which returns the next instruction in serial
program order, if we have accumulated enough memory responses to execute it
accurately, or returns
$\epsilon$ otherwise.  When $\commit$ returns an instruction, it also
returns the associated program counter plus the next program counter
we would advance to afterward.
 Furthermore, the instruction is
extended with any relevant response from memory (used only for load
instructions, obtained through $\upd\ld$) and with the new architectural state (\eg{} register
file) after execution.
\item $\retire(\rob)$, which informs the buffer that its $\commit$
instruction was executed successfully, so it is time to move on to the
next instruction.
\end{itemize}

@@

\inference[Compute]
{}
{\ioe{\text{P$_\text{so}$}}{s,\pc,\rob,\ppc}{s,\pc,\compute(\rob),\ppc}}

\inference[SpecLd]
{(\spc\ld,t,a) \in \get\spc\ld(\rob)}
{\io{\text{P$_\text{so}$}}{s,\pc,\rob,\ppc}{s,\pc,\upd\ld(\rob,t,v),\ppc}
{(\ld\req, a), (\ld\resp, v), \Spec}}

@@

\inference[Load]
{\s(c,a) \geq S}
{\io{$M_c$}
{\dt, \ch, \s, \dst, \wt, \dwt}
{\dt, \ch, \s, \dst, \wt, \dwt}{c,((\ld\req, a),(\ld\resp,v), \_)}}
\vspace{.1in}
\inference[Store]
{s(c,a) \geq M}
{\io{$M_c$}
{\dt, \ch, \s, \dst, \wt, \dwt}
{\dt[(c, a) \coloneqq v], \ch, \s, \dst, \wt, \dwt}{c,(\st,a,v)}}
